---
title: "mixed effects models"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(ggplot2)

```

## R Markdown

[source](http://www.bodowinter.com/tutorial/bw_LME_tutorial2.pdf bodo winter tutorial 2)

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r data}
pitch = c(233,204,242,130,112,142)
sex = c(rep("female",3),rep("male",3))
my.df = data.frame(sex,pitch)
print(my.df)
xmdl = lm(pitch ~ sex, my.df)
summary(xmdl)
```
```{r}
age = c(14,23,35,48,52,67)
pitch = c(252,244,240,233,212,204)
my.df = data.frame(age,pitch)
xmdl = lm(pitch ~ age, my.df)
summary(xmdl)

```
```{r}
politeness <- 
read.csv("http://www.bodowinter.com/tutorial/politeness_data.csv")
```

```{r}
summary(politeness)
```
## Including Plots

You can also embed plots, for example:

```{r attritude, echo=FALSE}
ggplot(politeness,aes(y=frequency, x=gender:attitude)) +geom_boxplot()

```
```{r}
politeness.model = lmer(frequency ~ attitude +
  (1|subject) +
  (1|scenario), data=politeness)
summary(politeness.model)

```

```{r}
politeness.model = lmer(frequency ~ attitude +
gender + (1|subject) +
(1|scenario), data=politeness)
summary(politeness.model)
```
Note that compared to our earlier model without the fixed effect gender, the
variation that’s associated with the random effect “subject” dropped considerably.
This is because the variation that’s due to gender was confounded with the
variation that’s due to subject

intercept now represents female pitch (informal???)
attitu

?why is scenario considered random effect rather than fixed effect?


```{r}
politeness.null = lmer(frequency ~ gender +
(1|subject) + (1|scenario), data=politeness,
REML=FALSE)
politeness.null = lmer(frequency ~ attitude + gender + 
(1|subject) + (1|scenario), data=politeness,
REML=FALSE)
anova(politeness.null,politeness.model)
```

